---
title: "p8105_hw1_cl4773.Rmd"
author: "Clement Li"
date: "2025-09-17"
output: github_document
---
### 1) Load the moderndive library, and use the following code to load the early_january_weather dataset:
```{r}
library(moderndive)
data("early_january_weather")
```

#### The early_january_weather dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. The variables include: time_hour: The hour of the day.
temp: Temperature in degrees Celsius. humid: Humidity percentage. Other variables include wind speed, pressure, etc. 

The mean temperature is `r round(mean(early_january_weather$temp, na.rm = TRUE))` degrees Celsius.

#### Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable (adding color = ... inside of aes in your ggplot code should help). Describe patterns that are apparent in this plot.

```{r}
library(ggplot2)

graph = ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(title = "Temperature vs Time of Day Colored by Humidity",
       x = "Time (Hours)",
       y = "Temperature (°C)",
       color = "Humidity (%)") +
  theme_minimal()

print(graph)
```

#### Export your scatterplot to your project directory using ggsave.
```{r}
ggsave(filename = "scatterplot.png", plot = graph, width = 7, height = 5)

```


### 2) Create a data frame comprised of: a random sample of size 10 from a standard Normal distribution, a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 10, a factor vector of length 10, with 3 different factor “levels”.
```{r}
library(tidyverse)
set.seed(100) 
sample_values = rnorm(10)  # standard Normal sample
logical_vector = sample_values > 0  # logical vector
character_vector = sample(letters[1:5], 10, replace = TRUE)  # random characters
factor_vector = factor(sample(c("low", "medium", "high"), 10, replace = TRUE))  # factor with 3 levels

# create df
df <- tibble(
  sample_values,
  logical_vector,
  character_vector,
  factor_vector
)
```

#### Try to take the mean of each variable in your dataframe. What works and what doesn’t? Hint: for now, to take the mean of a variable in a dataframe, you need to pull the variable out of the dataframe. Try loading the tidyverse and using the pull function.
```{r}
mean(df %>% pull(sample_values))       # Works: numeric
mean(df %>% pull(logical_vector))      # Works: TRUE = 1, FALSE = 0
mean(df %>% pull(character_vector))    # Error: characters can't be averaged
mean(df %>% pull(factor_vector))       # Error: factors need to be numeric
```

#### In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?
```{r, results='hide', warning=FALSE}
as.numeric(df$logical_vector)      # TRUE becomes 1, FALSE becomes 0
as.numeric(df$character_vector)    # Converts to NA with warning 
as.numeric(df$factor_vector)       # Converts to underlying integer codes of factor levels
```

Only numeric and logical vectors give meaningful means. Character vectors cannot be averaged, and factors must be carefully converted if you want to compute summaries like means




















